
\clearpage

# Hippocampus and photoperiod

This section is guided by the paper

* N. A. Majrashi, "Sex differences in the association of photoperiod with hippocampal
subfield volumes in older adults: A cross-sectional study in the UK Biobank cohort",
_Brain Behavior_, June 2020, 10(6):e01593.

## Visualize subject home location and photoperiod

The authors used acquisition site instead of location of residence:

> Participants lived in approximately equal proportions north and south of the scanner
> and a mean distance of 31.1 km north or south of it (range 0.4–289 km). Therefore,
> location of residence would have a negligible effect on photoperiod and we used the
> photoperiod at the scanner location for all participants.

Here, we use residence location.

```{r photoperiod-plot, message = FALSE, results = 'asis', echo = FALSE, eval = TRUE}
library( ggmap )
library( geosphere )

imagingSite <- data.frame(
                Location = c( "Cheadle", "Reading", "Newcastle", "Bristol" ),
                Coding = c( 11025, 11026, 11027, 11028 ),
                Longitude = c( -2.21, -0.978, -1.62, -2.59 ),
                Latitude = c( 53.4, 51.5, 55.0, 51.5 ) )
useImagingSite <- FALSE

baseDirectory <- "../Data/ukbiobank/"

ukbAll <- read.csv( paste0( baseDirectory, "ukb50844_imaging_only_Brian.csv" ) )
ukbAll$photoPeriod <- daylength( ukbAll$Latitude, ukbAll$f.53.2.0 )

if( useImagingSite )
  {
  for( i in seq.int( nrow( imagingSite ) ) )
    {
    ukbAll$Latitude[data$f.54.2.0 == imagingSite$Coding[i]] <- imagingSite$Latitude[i]
    ukbAll$Longitude[data$f.54.2.0 == imagingSite$Coding[i]] <- imagingSite$Longitude[i]
    }
  } else {
  ukbAll <- ukbAll[which( ! is.na( ukbAll$Longitude ) ),]
  }

numberOfSubjectsToPlot <- 5000
subjectIndices <- sample.int( nrow( ukbAll ), size = numberOfSubjectsToPlot )
plotData <- ukbAll[subjectIndices,]

ukMapFile <- "../Scripts/ukMap.RData"
if( file.exists( ukMapFile ) )
  {
  load( ukMapFile )
  } else {
  ukMap <- get_map( location = "Great Britain", source = "google", maptype = "terrain", zoom = 5 )
  }

ukPlot <- ggmap( ukMap ) +
             geom_point( aes( x = Longitude, y = Latitude, color = photoPeriod ), data = plotData, alpha = 0.3, size = 1 ) +
             scale_colour_gradientn( colours = rev( c( 'gold', 'orange', 'red', 'darkred', 'red4', 'black'  ) ) ) +
             xlab( "Longitude" ) +
             ylab( "Latitude" ) +
             ggtitle( "Home location" )
print( ukPlot )
# ggsave( paste0( baseDirectory, "ukbbMap.pdf" ), ukPlot, units = "in", width = 10, height = 10 )
```


## Correlations on whole hippocampal volumes

```{r photoperiod-partial-correlations, message = FALSE, results = 'asis', echo = FALSE, eval = TRUE}

##########################
#
# look at hippocampal volumes
#

idpData <- loadIdpData( whichData = c( "fsl_other",
                                       "fsl_fast",
                                       "freesurfer_aseg",
                                       "freesurfer_hippocampus_mapped2deepflash",
                                       "ants_deep_flash",
                                       "hippmapper" ),
                        completeCases = TRUE,
                        commonSubjects = TRUE )

fsl_other <- data.frame( Subject = idpData$fsl_other$Subject,
                         right.hippocampus = idpData$fsl_other$f.25020,
                         left.hippocampus = idpData$fsl_other$f.25019 )
fsl_other <- fsl_other[which( fsl_other$Subject %in% ukbAll$f.eid ),]

fsl_fast <- data.frame( Subject = idpData$fsl_fast$Subject,
                        right.hippocampus = idpData$fsl_fast$f.25887,
                        left.hippocampus = idpData$fsl_fast$f.25886 )
fsl_fast <- fsl_fast[which( fsl_fast$Subject %in% ukbAll$f.eid ),]

freesurfer_aseg <- data.frame( Subject = idpData$freesurfer_aseg$Subject,
                               right.hippocampus = idpData$freesurfer_aseg$f.26593,
                               left.hippocampus = idpData$freesurfer_aseg$f.26562 )
freesurfer_aseg <- freesurfer_aseg[which( freesurfer_aseg$Subject %in% ukbAll$f.eid ),]

freesurfer_hippocampus <- data.frame( Subject = idpData$freesurfer_hippocampus_mapped2deepflash$Subject,
                                      right.hippocampus = idpData$freesurfer_hippocampus_mapped2deepflash$right.hippocampus,
                                      left.hippocampus = idpData$freesurfer_hippocampus_mapped2deepflash$left.hippocampus )
freesurfer_hippocampus <- freesurfer_hippocampus[which( freesurfer_hippocampus$Subject %in% ukbAll$f.eid ),]

ants_deep_flash <- data.frame( Subject = idpData$ants_deep_flash$Subject,
                               right.hippocampus = idpData$ants_deep_flash$right.hippocampus,
                               left.hippocampus = idpData$ants_deep_flash$left.hippocampus )
ants_deep_flash <- ants_deep_flash[which( ants_deep_flash$Subject %in% ukbAll$f.eid ),]

ants_deep_flash_mtl <- data.frame( Subject = idpData$ants_deep_flash$Subject,
                                   right.hippocampus = idpData$ants_deep_flash$right.MTL,
                                   left.hippocampus = idpData$ants_deep_flash$left.MTL )
ants_deep_flash_mtl <- ants_deep_flash_mtl[which( ants_deep_flash_mtl$Subject %in% ukbAll$f.eid ),]

ants_deep_flash_other <- data.frame( Subject = idpData$ants_deep_flash$Subject,
                                     right.hippocampus = idpData$ants_deep_flash$right.extra.hippocampal,
                                     left.hippocampus = idpData$ants_deep_flash$left.extra.hippocampal )
ants_deep_flash_other <- ants_deep_flash_other[which( ants_deep_flash_other$Subject %in% ukbAll$f.eid ),]

hippmapper <- idpData$hippmapper[which( idpData$hippmapper$Subject %in% ukbAll$f.eid ),]

ukbAll <- ukbAll[which( ukbAll$f.eid %in% hippmapper$Subject ),]

hippocampalPipelines <- list()
hippocampalPipelines[[1]] <- fsl_fast
hippocampalPipelines[[2]] <- fsl_other
hippocampalPipelines[[3]] <- freesurfer_aseg
hippocampalPipelines[[4]] <- freesurfer_hippocampus
hippocampalPipelines[[5]] <- hippmapper
hippocampalPipelines[[6]] <- ants_deep_flash
# hippocampalPipelines[[7]] <- ants_deep_flash_mtl
# hippocampalPipelines[[8]] <- ants_deep_flash_other
# hippocampalPipelineNames <- c( "FSLfast", "FSLother", "FSaseg", "FShipp", "Hippmapper", "DeepFlash", "DeepFlashMtl", "DeepFlashOther" )
hippocampalPipelineNames <- c( "FSLfast", "FSLother", "FSaseg", "FShipp", "Hippmapper", "DeepFlash" )

ppDf <- data.frame( HippocampalVolume = rep( NA, length( ukbAll$photoPeriod ) ),
                    PhotoPeriod = ukbAll$photoPeriod,
                    EuclideanDistanceToCoast = ukbAll$f.24508.0.0,
                    Age = ukbAll$f.21003.2.0,
                    Sex = ukbAll$f.22001.0.0,
                    Site = ukbAll$f.54.2.0,
                    TotalBrainVolume = ukbAll$f.25010.2.0 )

for( i in seq.int( length( hippocampalPipelines ) ) )
  {
  cat( "* Partial correlation (Volume ~ PhotoPeriod) for ", hippocampalPipelineNames[i], "\n\n" )

  hippoFormula <-  as.formula( "HippocampalVolume ~ Age + TotalBrainVolume + PhotoPeriod + Sex" )

  hippoLm <- list()
  for( hemi in c( "Left", "Right" ) )
    {
    if( hemi == "Left" )
      {
      ppDf$HippocampalVolume <- hippocampalPipelines[[i]]$left.hippocampus
      hippoLm[[1]] <- lm( hippoFormula, data = ppDf )
      } else {
      ppDf$HippocampalVolume <- hippocampalPipelines[[i]]$right.hippocampus
      hippoLm[[2]] <- lm( hippoFormula, data = ppDf )
      }

    volumeResiduals <- lm( as.formula( "HippocampalVolume ~ Age + TotalBrainVolume + Sex" ), data = ppDf )$residuals
    photoPeriodResiduals <- lm( as.formula( "PhotoPeriod ~ Age + TotalBrainVolume + Sex" ), data = ppDf )$residuals

    correlationValue <- cor( photoPeriodResiduals, volumeResiduals )
    correlationP <- cor.test( photoPeriodResiduals, volumeResiduals )$p.value

    cat( "    *", hemi, ":", correlationValue, "( p.value =", correlationP, ")\n\n" )
    }
  }

```

\clearpage

## Regression on whole hippocampal volumes

```{r photoperiod-regressions, message = FALSE, results = 'asis', echo = FALSE, eval = TRUE}

cat( "**Formula:  **\n" )
print( hippoFormula )
cat( "\n\n" )

for( i in seq.int( length( hippocampalPipelines ) ) )
  {
  hippoLm <- list()
  for( hemi in c( "Left", "Right" ) )
    {
    if( hemi == "Left" )
      {
      ppDf$HippocampalVolume <- hippocampalPipelines[[i]]$left.hippocampus
      hippoLm[[1]] <- lm( hippoFormula, data = ppDf )
      } else {
      ppDf$HippocampalVolume <- hippocampalPipelines[[i]]$right.hippocampus
      hippoLm[[2]] <- lm( hippoFormula, data = ppDf )
      }
    }
  stargazer::stargazer( hippoLm[[1]], hippoLm[[2]],
      type = "latex", header = FALSE,
      column.labels = brainPipelineNames[c( "Left", "Right" )],
      label = paste0( 'table::photoperiod', i ),
      title = hippocampalPipelineNames[i] )
  }

```